Uncovering underlying musical content in the temporal domain \\ \small Leveraging deep learning, inductive bias, and aural music skills to learn embedding spaces from raw waveforms with applications to MIR downstream tasks

### 2023-05-02 introduction/Subsections/Motivation.tex
Updates from Overleaf

Personal, academic, and industrial motivations drive this thesis. It is conducted in collaboration with \textit{Epidemic Sound AB (ES)} \cite{EpidemicSite}, a prominent Swedish company that provides a global library of over 40,000+ royalty-free soundtrack music and 90,000+ sound effects, and the \textit{Music Technology Group (MTG)} at Universitat Pompeu Fabra in Barcelona, Spain.

@@

As a music enthusiast, professional musician, and scholar, I am passionately committed to exploring the foundational elements of musical composition, investigating the diverse techniques employed in creating musical works, and deciphering the intricate relationships among them. This passion fuels my quest for knowledge on extracting valuable information embedded within music in any domain, contributing to developing more sophisticated and effective music-related technologies and solutions in the industrial landscape.

I advocate for exposure as the primary method of learning music, emphasizing the importance of immersing oneself in various musical experiences to foster a holistic understanding of music as an art form. Individuals can cultivate well-rounded musical expertise by engaging with different styles, genres, techniques, and learning approaches, ultimately promoting creativity and experimentation in both personal and professional domains.

@@

\begin{figure}[ht]
\includegraphics[clip,width=\columnwidth]{figures/schenkerian analysis/SchubertOp4no3.png}% 
\caption{Small excerpt of \textit{Wandrers Nachtlied, Op. 4, D. 224} by Franz Schubert. We can see the passage's original score, the schenkerian unfolding of the melody, the chord degrees analysis, and the tonal function.}
\label{fig:Wandrers Nachtlied, Op. 4, D. 224}
\end{figure}

>>>>

Personal, academic, and industrial motivations drive this thesis.

This research is conducted in collaboration with \textit{Epidemic Sound AB (ES)} \cite{EpidemicSite}, a prominent Swedish company that provides a global library of over 40,000+ royalty-free soundtrack music and 90,000+ sound effects, and the \textit{Music Technology Group (MTG)} at Universitat Pompeu Fabra in Barcelona, Spain.

@@

As a music enthusiast, professional musician, and scholar, I am passionately committed to exploring the foundational elements of musical composition, investigating the diverse techniques employed in creating musical works, and deciphering the intricate relationships among them. This passion fuels my quest for knowledge on extracting valuable information embedded within music in any domain, being my main interest in the audio and symbolic domain, contributing to developing more sophisticated and effective music-related technologies and solutions in the industrial landscape.

I advocate for exposure as the primary method of learning music, emphasizing the importance of immersing oneself in various musical experiences to foster a holistic understanding of music as an art form. Individuals can cultivate well-rounded musical expertise by engaging with different styles, genres, techniques, and learning approaches, ultimately promoting creativity and experimentation.

@@

\begin{figure}[ht]
\includegraphics[clip,width=\columnwidth]{figures/schenkerian analysis/SchubertOp4no3.png}% 
\caption{\small{Small excerpt of \textit{Wandrers Nachtlied, Op. 4, D. 224} by Franz Schubert. We can see the passage's original score, the schenkerian unfolding of the melody, the chord degrees analysis, and the tonal function.}}
\label{fig:Wandrers Nachtlied, Op. 4, D. 224}
\end{figure}

### 2023-05-02 introduction/Subsections/Objectives.tex
Updates from Overleaf

Neural networks have recently facilitated the development of low-dimensional musical embeddings that capture essential musical features, enabling the computation of content-based similarity between musical passages \cite{KimLEARNINGLOSS}\cite{Hung2022Feature-informedClassification}. An embedding function $f(x) \in \mathbb{R}^d$ maps a given course $x$ to a vector in the real space $\mathbb{R}^d$, allowing the measurement of similarity between two musical passages using Euclidean distance, as denoted by the formula $D_{i,j} = ||f(x_i) - f(x_j)||^2$. The success of such systems hinges on the level of musical abstraction and the embedding space's capacity to interpret music in a manner consistent with human perception. As a result, a suitable embedding space is vital for producing output that aligns with human expectations.

@@

@@

For a given music piece $M$, we aim to learn a single embedding, $e(M)$, that captures both the underlying content (melody, harmony, critical, holistic elements) and production aspects. The embedding space should be structured to cluster similar content and production features together, effectively separating the two aspects. Formally, our objective is to learn the embedding function $e(M)$ such that:

@@

We propose learning a single embedding $e(M)$ that captures content and production features, with the embedding space structured to cluster similar aspects together, effectively distinguishing between the two.

@@

Additionally, this research contends that MIR studies should transition from a solely technical-centered approach to a more balanced and comprehensive method that integrates other pertinent domains such as musicology and music theory. This shift will enable a better understanding of the interplay between technical and musical aspects of music, enhancing our comprehension of music signals and music itself.

>>>>

@@

Neural networks have recently facilitated the development of low-dimensional musical embeddings that capture essential musical features, enabling the computation of content-based similarity between musical passages \cite{Kim2021LearningLoss}\cite{Hung2022Feature-informedClassification}. An embedding function $f(x) \in \mathbb{R}^d$ maps a given course $x$ to a vector in the real space $\mathbb{R}^d$, allowing the measurement of similarity between two musical passages using Euclidean distance, as denoted by the formula $D_{i,j} = ||f(x_i) - f(x_j)||^2$. The success of such systems hinges on the level of musical abstraction and the embedding space's capacity to interpret music in a manner consistent with human perception. As a result, a suitable embedding space is vital for producing output that aligns with human expectations.

@@

For a given music piece $M$, we aim to learn a single embedding, $e(M)$, that captures the underlying content (melody, harmony, critical, holistic elements) and production aspects. The embedding space should be structured to cluster similar content and production features, effectively separating the two elements. Formally, our objective is to learn the embedding function $e(M)$ such that:

@@

We propose learning a single embedding $e(M)$ that captures musical high-level content while keeping agnostic to production features. The embedding space is structured to cluster similar aspects together, effectively distinguishing between the two.

@@

Additionally, this research contends that MIR studies should transition from a solely technical-centered approach to a more balanced and comprehensive method that integrates other pertinent domains, such as musicology and music theory. This shift will enable a better understanding of the interplay between technical and musical aspects of music, enhancing our comprehension of music signals and the music itself.

### 2023-05-02 methods/subsections/machine learning.tex
Updates from Overleaf

\section{Artificial Neural Networks (ANNs)}
A neural network, formally known as an artificial neural network (ANN), is a computational model inspired by the structure and functionality of biological neural networks. It is a mathematical framework designed to process information and perform tasks like pattern recognition, classification, and regression by mimicking how biological neurons and synapses operate. ANNs comprise interconnected layers of artificial neurons or nodes, each performing a mathematical transformation on the input data to generate an output. The connections between these neurons, often called "weights," are adjusted during learning to optimize the network's performance on a given task.

@@

\item \textbf{Convolutional layers:} These layers apply a convolution operation to the input data. The convolutional layers are designed to effectively learn local patterns and features in the audio data \footnote{For a visual explanation and further understanding of the operation convolution, see this video: \url{https://youtu.be/KuXjwB4LzSA}}.

@@

$$(I * K)_{kl} = \sum_{i=1}^{p} \sum_{j=1}^{q} i_{k+i-1, l+j-1}k_{ij}$$

>>>>

\section{Artificial Neural Networks (ANNs)}
A neural network, formally known as an Artificial Neural Network (ANN), is a computational model inspired by the structure and functionality of biological neural networks. It is a mathematical framework designed to process information and perform tasks like pattern recognition, classification, and regression by mimicking how biological neurons and synapses operate. ANNs comprise interconnected layers of artificial neurons or nodes, each performing a mathematical transformation on the input data to generate an output. The connections between these neurons, often called "weights," are adjusted during learning to optimize the network's performance on a given task.

@@

\item \textbf{Convolutional layers:} These layers apply a convolution operation to the input data. The convolutional layers are designed to effectively learn local patterns and features in the audio data \footnote{For a visual explanation and further understanding of the convolution operation, please refer to this video: \url{https://youtu.be/KuXjwB4LzSA}}.

@@

\begin{equation}
    (I * K)_{kl} = \sum_{i=1}^{p} \sum_{j=1}^{q} i_{k+i-1, l+j-1}k_{ij}
\end{equation}
